---
title: "Global HIV/AIDS Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load the libraries
library("dplyr")
library("plotly")
library("ggplot2")
library("leaflet")
library("tidyverse")
library("knitr")

# Loading data
hiv_df <- read.csv("data/art_coverage_by_country_clean.csv",
  stringsAsFactors = FALSE
)

# Data for country coordinates
country_coordinates <- read.csv("data/world_country_coordinate.csv",
  stringsAsFactors = FALSE
)
```

## Introduction 
The data was collected and generated by WHO and UNESCO from their public records
on the number of people living with HIV/AIDS around the world. Devakumar cleaned
that data from WHO and UNESCO and put the data into a CSV file that draws a focus
on treatment statistics for those living with HIV/AIDS by country.There are 170
observations (rows) in the HIV/AIDS dataset and 10 features (columns) in the
HIV/AIDS dataset. With this dataset we are going to try and view the extent of
treatment coverage is for HIV/AIDS in countries globally, ART treatment's impact
on the number of people living with HIV, and compare the amount of people living
with HIV by region. We added a second dataset from <https://www.kaggle.com/paultimothymooney/latitude-and-longitude-for-every-country-and-state>
to add longitude and latitude for each country. 

## Summary Information

```{r summary, echo=FALSE, warning=FALSE}
source("scripts/summaryinfo.R")
summary_info <- get_summary_info(hiv_df)
```

The most common treatment for HIV around the world is antiretroviral therapy (ART).
Among the `r summary_info$num_countries_hiv` countries that are estimated to
have people living with HIV based on the WHO and UNESCO data,
`r summary_info$num_countries_art` have a reported number of people with ART
coverage. The ART treatment is effective at allowing people with HIV to live
longer and have better overall health than they would otherwise, however it is
not a cure for HIV. Another issue is that the amount of ART coverage globally
is not distributed equitably. For example, the WHO region that has the highest
proportion of people reportedly receiving ART coverage to the estimated number
of people living with HIV is `r summary_info$max_region_cov_prop_region`,
with a coverage proportion of `r summary_info$max_region_cov_prop`,
which is far higher than the coverage proportion of `r summary_info$min_region_cov_prop`,
in the `r summary_info$min_region_cov_prop_region`. This displays that HIV is
still a huge global health problem in some regions of the world.
Despite the `r summary_info$min_region_cov_prop_region` having the lowest ART
coverage proportion, the region in which HIV cases themselves are most prevalent
is Africa. Specifically, `r summary_info$max_cases_country` has the most
estimated people living with HIV of any country on record,
with `r summary_info$max_cases` cases, and specific countries within Africa have
an individual coverage proportion lower than any world region.
`r summary_info$min_coverage_prop_country` for example, has the lowest coverage
proportion of any country in the world with a `r summary_info$min_coverage_prop`.
Due to the still massive amount of HIV cases in the world and the dismally low
ART coverage in some areas, expanding ART coverage and making it more accessible
is clearly something that needs to continue to be a priority globally.
The following table, map, and charts will further cement this point.

## Aggregate Table
The table below is organized by World Health Organization (WHO) regions. It
displays the total cases of HIV, total ART coverage, and the proportion of ART
coverage to people living with HIV in each region. The regions are ordered by
greatest to least proportion of ART coverage to those who have HIV. This table
is beneficial as it allows us to view which regions have effective ART coverage
for those living with HIV, and which countries need more help with providing
that coverage. 

The table below is organized by World Health Organization (WHO) regions. It
displays the total cases of HIV, total ART coverage, and the proportion of ART
coverage to people living with HIV in each region. The regions are ordered by
greatest to least proportion of ART coverage to those who have HIV. This table
is beneficial as it allows us to view which regions have effective ART coverage
for those living with HIV, and which countries need more help with providing
that coverage.

```{r aggregate-table, echo=FALSE, warning=FALSE}
source("scripts/tablesummary.R")
kable(summary_table(hiv_df))
```


## Charts

```{r firstchart, echo=FALSE, warning=FALSE}
source("scripts/firstchart.R")
geographical_chart(hiv_df, country_coordinates)
```

This is an interactive map that plots the geographic location for each country.
The size of the marker is determined by the number of people who have
received ART. When the marker is clicked it provides the country name and the
exact number of people who have received ART. There is also a color range that
goes from red to blue to indicate how many people are receiving ART.We included
this chart to visually see the locations of the countries and compare the number
of people receiving ART in each country based on the size of the circle.
We can see that Africa has is the region where most of the data for number of
people receiving ART are from. Visually, South Africa has the highest number
of people receiving ART based on the size and color.

```{r secondchart.R, echo=FALSE}
source("scripts/secondchart.R")
art_treated_vs_sick(hiv_df)
```

This scatter plot displays the trend of of how the estimated number people
living with HIV is impacted as the number of people receiving art increases.
There is an overall linear, positive trend with variation. This positive trend
implies that as the estimated amount of people receiving ART increases,
the estimated amount of people living with HIV also increases.This increase in
people living with HIV is due to the treatment being effective, and allowing for
those living with HIV to live longer lives.

```{r thirdchart, echo=FALSE, warning=FALSE}
source("scripts/thirdchart.R")
plot_hiv_per_region(hiv_df)
```

This chart was intended to show the relationship between the number of HIV cases
and the region of the world. The chart's goal was to help us figure out what 
region has been impacted the most by the HIV crisis. It is clear from looking at 
the bar chart that the African region has been hit the hardest, with 25.5 
million estimated cases. It has the highest total of estimated HIV cases than 
any other region. In fact, it has more cases than every other region put 
together. The next highest region is the America's with 2.3 million estimated 
cases. The Eastern Mediterranean region has the lowest amount of cases, with 
380 thousand cases. Most of the world's focus in dealing with the HIV crisis 
should be centered on helping Africa. 

